---
title: "Coursera/Project2"
author: "Rahul"
date: "20 December 2017"
output: 
  html_document: 
    keep_md: yes
    self_contained: no
---
Synopsis
==========

In this report we will analysis the NOAA Storm Events data. The data collected from year 1950 to 2011, 
with 902297 records. Fatalities, injuries, and property damage (in dollars) are totalled over that time.
Data will be cleansed, re-orderd and analyised in order to identified

Data Processing

```{r,cache= TRUE}
url = "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(url, dest = "storm.bz2")
stormData<-read.csv(bzfile("storm.bz2"))
head(stormData, 3)
```

Subsetting and summarised require Data 

```{r}
New_stormData<- stormData[,c("EVTYPE", "FATALITIES", "INJURIES", "PROPDMG","PROPDMGEXP","CROPDMG","CROPDMGEXP")]

summary(New_stormData)
```

assigning new column name

```{r}
New_stormData$EVTYPE <- 
        gsub("^HEAT$", " EXCESSIVE HEAT", New_stormData$EVTYPE)
New_stormData$EVTYPE <- 
        gsub("^TSTM WIND$", "THUNDERSTORM WIND", New_stormData$EVTYPE)
New_stormData$EVTYPE <- 
        gsub("^THUNDERSTORM WIND$", "THUNDERSTORM WINDS", New_stormData$EVTYPE)
```

The purpose of the following code is to aggregate the data in order to prepare it for the results. Data will be aggregated (SUM) on Event Type level, the TOP 10 events will be then indentified.

```{r}
agg_fatData <-
        aggregate(
                New_stormData$FATALITIES, 
                by=list(New_stormData$EVTYPE), FUN=sum, na.rm=TRUE)
colnames(agg_fatData) <- c("event.type", "fatality.total")
fatalities.sorted <- 
        agg_fatData[order(-agg_fatData$fatality.total),] 
top.fatalities <- fatalities.sorted[1:10,]
top.fatalities$event.type <- 
        factor(
                top.fatalities$event.type, levels=top.fatalities$event.type, 
                ordered=TRUE)
```

Ploting graph for Top 10  causes of Fatalities

In this case Tornado are most harmful events for population.

```{r}
library(ggplot2)
ggplot(data=top.fatalities, aes(x=event.type, y=fatality.total)) + 
        geom_bar(stat="identity",fill= "blue") + xlab("Event type") + ylab("Total fatalities") + 
        ggtitle("Fatalities By Event Type")+
        theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


Aggregate data on injuries to find which events are the top 10 causes of injuries

```{r}
agg_injData <-
        aggregate(
                New_stormData$INJURIES, 
                by=list(New_stormData$EVTYPE), FUN=sum, na.rm=TRUE)
colnames(agg_injData) <- c("event.type", "injury.total")
injuries.sorted <- agg_injData[order(-agg_injData$injury.total),] 
top.injuries <- injuries.sorted[1:10,]
top.injuries$event.type <- 
        factor(
                top.injuries$event.type, levels=top.injuries$event.type, 
                ordered=TRUE)
```                

Ploting graph for Top 10  causes of injuries

```{r}
ggplot(data=top.injuries, aes(x=event.type, y=injury.total)) + 
        geom_bar(stat="identity") + xlab("Event type") + ylab("Total injuries") + 
        ggtitle("Injuries By Event Type") +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Aggregate data on Property Damage to find which events are the top 10 causes of Property Damage.

```{r}
agg_prodamData <-
        aggregate(
                New_stormData$PROPDMG, 
                by=list(New_stormData$EVTYPE), FUN=sum, na.rm=TRUE)
colnames(agg_prodamData) <- c("event.type", "prop.dmg.total")
prop.dmg.sorted <- agg_prodamData[order(-agg_prodamData$prop.dmg.total),] 
top.prop.dmg <- prop.dmg.sorted[1:10,]
top.prop.dmg$event.type <- 
        factor(
                top.prop.dmg$event.type, levels=top.prop.dmg$event.type, 
                ordered=TRUE)
```                


Ploting graph for Top 10  causes of Property Damage.
```{r}
ggplot(data=top.prop.dmg, aes(x=event.type, y=prop.dmg.total)) + 
        geom_bar(stat="identity") + xlab("Event type") + 
        ylab("Total property damage") +  ggtitle("Property Damage By Event Type") + 
        theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Aggregate data on crop Damage to find which events are the top 10 causes of crop Damage.

```{r}
agg_cropData <-
        aggregate(
                New_stormData$CROPDMG, 
                by=list(New_stormData$EVTYPE), FUN=sum, na.rm=TRUE)
colnames(agg_cropData) <- c("event.type", "crop.dmg.total")
crop.dmg.sorted <- agg_cropData[order(-agg_cropData$crop.dmg.total),] 
top.crop.dmg <- crop.dmg.sorted[1:10,]
top.crop.dmg$event.type <- 
        factor(
                top.crop.dmg$event.type, levels=top.crop.dmg$event.type, 
                ordered=TRUE)
```                

Ploting graph for Top 10  causes of crop Damage.
This graph showa hail is most charmful cause for Crops

```{r}
ggplot(data=top.crop.dmg, aes(x=event.type, y=crop.dmg.total)) + 
        geom_bar(stat="identity") + xlab("Event type") + 
        ylab("Total crop damage") +  ggtitle("Property Damage By Event Type") + 
        theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


Tornados caused the maximum number of fatalities and injuries